---
title: "A Quick Introduction to iNEXT.3D via Examples"
author: "Anne Chao"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    vignette: >
      %\VignetteIndexEntry{A Quick Introduction to iNEXT.3D via Examples}
      %\VignetteEngine{knitr::rmarkdown}
      %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "", 
                      fig.retina=2,
                      fig.align='center',
                      fig.width = 7, fig.height = 5,
                      warning = FALSE, message = FALSE)
options("width"=200)

if("gridExtra" %in% rownames(installed.packages()) == FALSE) {install.packages("gridExtra")}
devtools::load_all(".")
library(iNEXT.3D)
library(ggplot2)
data("beetles.abu")
```


```{r eval=FALSE}
## install iNEXT.3D package from CRAN
# install.packages("iNEXT.3D")
## install the latest version from github
install.packages('devtools')
library(devtools)
install_github('AnneChao/iNEXT.3D')
## import packages
library(iNEXT.3D)
library(ggplot2)
```


In this document, we provide a quick introduction demonstrating how to run the package `iNEXT.3D`(iNterpolation and EXTrapolation in three Dimensions). `iNEXT.3D` has several main function: `iNEXT3D`, `ggiNEXT3D`, `obs3D`, `asy3D`, `ggasy3D`, and `DataInfo3D.`

`iNEXT3D` focuses on three measures of Hill numbers of order q: species richness (`q = 0`), Shannon diversity (`q = 1`, the exponential of Shannon entropy) and Simpson diversity (`q = 2`, the inverse of Simpson concentration) among three classes: taxonomic diversity ('TD'), phylogenetic diversity ('PD'), functional diversity ('FD'). For each diversity measure, `iNEXT.3D` uses the observed sample of abundance or incidence data (called the “reference sample”) to compute diversity estimates and the associated confidence intervals for the following two types of rarefaction and extrapolation (R/E): 

1. Sample‐size‐based R/E sampling curves:`iNEXT3D` computes diversity estimates for rarefied and extrapolated samples up to an appropriate size. This type of sampling curve plots the diversity estimates with respect to sample size.     
2. Coverage‐based R/E sampling curves: `iNEXT3D` computes diversity estimates for rarefied and extrapolated samples with sample completeness (as measured by sample coverage) up to an appropriate coverage. This type of sampling curve plots the diversity estimates with respect to sample coverage. 

`iNEXT3D` also plots the above two types of sampling curves and a sample completeness curve. The sample completeness curve provides a bridge between these two types of curves.

## MAIN FUNCTION: iNEXT3D()

We first describe the main function `iNEXT3D()` with default arguments: 

```{r eval=FALSE}
iNEXT3D(data, diversity = 'TD', q = c(0,1,2), datatype = "abundance", 
        size = NULL, endpoint = NULL, knots = 40, nboot = 50, conf = 0.95, nT = NULL, 
        PDtree = NULL, PDreftime = NULL, PDtype = 'meanPD', 
        FDdistM, FDtype = 'AUC', FDtau = NULL)
```

The arguments of this function are briefly described below, and will be explained in more details by illustrative examples in later text. This main function computes diversity estimates of order q, the sample coverage estimates and related statistics for K (if `knots=K`) evenly‐spaced knots (sample sizes) between size 1 and the `endpoint`, where the endpoint is described below. Each knot represents a particular sample size for which diversity estimates will be calculated. By default, endpoint = double the reference sample size (total sample size for abundance data; total sampling units for incidence data). For example, if `endpoint = 10`, `knot = 4`, diversity estimates will be computed for a sequence of samples with sizes (1, 4, 7, 10).  
<table style="width:100%;">
<colgroup>
<col width="20%">
<col width="80%">
</colgroup>
<thead>
<tr class="header">
<th align="center">Argument</th>
<th align="left">Description</th>
</tr>
</thead>
<tbody>
<tr class="odd">
<td align="center"><code>data</code></td>
<td align="left">
(a) For </code>datatype = "abundance"</code>, data can be input as a vector of species abundances (for a single assemblage), matrix/data.frame (species by assemblages), or a list of species abundance vectors. \cr
(b) For </code>datatype = "incidence_freq"</code>, data can be input as a vector of incidence frequencies (for a single assemblage), matrix/data.frame (species by assemblages), or a list of incidence frequencies; the first entry in all types of input must be the number of sampling units in each assemblage. \cr
(c) For </code>datatype = "incidence_raw"</code>, data can be input as a list of matrix/data.frame (species by sampling units); data can also be input as a matrix/data.frame by merging all sampling units across assemblages based on species identity; in this case, the number of sampling units (nT, see below) must be input. </td>
</tr>
<tr class="even">
<td align="center"><code>diversity</code></td>
<td align="left">selection of diversity type: </code>'TD'</code> = Taxonomic diversity, </code>'PD'</code> = Phylogenetic diversity, and </code>'FD'</code> = Functional diversity.</td>
</tr>
<tr class="odd">
<td align="center"><code>q</code></td>
<td align="left">a numerical vector specifying the diversity orders. Default is c(0, 1, 2).</td>
</tr>
<tr class="even">
<td align="center"><code>datatype</code></td>
<td align="left">data type of input data: individual-based abundance data (</code>datatype = "abundance"</code>), sampling-unit-based incidence frequencies data (</code>datatype = "incidence_freq"</code>), or species by sampling-units incidence matrix (</code>datatype = "incidence_raw"</code>) with all entries being 0 (non-detection) or 1 (detection).</td>
</tr>
<tr class="odd">
<td align="center"><code>size</code></td>
<td align="left">an integer vector of sample sizes (number of individuals or sampling units) for which diversity estimates will be computed.
If NULL, then diversity estimates will be computed for those sample sizes determined by the specified/default </code>endpoint</code> and </code>knots</code>.</td>
</tr>
<tr class="even">
<td align="center"><code>endpoint</code></td>
<td align="left">an integer specifying the sample size that is the </code>endpoint</code> for rarefaction/extrapolation.
If NULL, then </code>endpoint</code> </code>=</code> double reference sample size.</td>
</tr>
<tr class="odd">
<td align="center"><code>knots</code></td>
<td align="left">an integer specifying the number of equally-spaced </code>knots</code> (say K, default is 40) between size 1 and the </code>endpoint</code>;each knot represents a particular sample size for which diversity estimate will be calculated.
If the </code>endpoint</code> is smaller than the reference sample size, then </code>iNEXT3D()</code> computes only the rarefaction esimates for approximately K evenly spaced </code>knots</code>. 
If the </code>endpoint</code> is larger than the reference sample size, then </code>iNEXT3D()</code> computes rarefaction estimates for approximately K/2 evenly spaced </code>knots</code> between sample size 1 and the reference sample size, and computes extrapolation estimates for approximately K/2 evenly spaced </code>knots</code> between the reference sample size and the </code>endpoint</code>.</td>
</tr>
<tr class="even">
<td align="center"><code>nboot</code></td>
<td align="left">a positive integer specifying the number of bootstrap replications when assessing sampling uncertainty and constructing confidence intervals. Enter 0 to skip the bootstrap procedures. Default is 50.</td>
</tr>
<tr class="odd">
<td align="center"><code>conf</code></td>
<td align="left">a positive number < 1 specifying the level of confidence interval. Default is 0.95.</td>
</tr>
<tr class="even">
<td align="center"><code>nT</code></td>
<td align="left">(required only when </code>datatype = "incidence_raw"</code> and input data is matrix/data.frame) a vector of nonnegative integers specifying the number of sampling units in each assemblage. If assemblage names are not specified, then assemblages are automatically named as "assemblage1", "assemblage2",..., etc. </td>
</tr>
<tr class="odd">
<td align="center"><code>PDtree</code></td>
<td align="left">(required only when </code>diversity = "PD"</code>), a phylogenetic tree in Newick format for all observed species in the pooled assemblage.</td>
</tr>
<tr class="even">
<td align="center"><code>nT</code></td>
<td align="left">an vector needed for class <code>'PD'</code> if the datatype belong to <code>"incidence_raw"</code> and data form belog to <code>matrix</code> or <code>data.frame</code>. The vector specifying the sampling unit of each assemblage.</td>
</tr>
<tr class="odd">
<td align="center"><code>PDreftime</code></td>
<td align="left">(required only when </code>diversity = "PD"</code>), a vector of numerical values specifying reference times for PD. Default is </code>NULL</code> (i.e., the age of the root of PDtree).</td>
</tr>
<tr class="even">
<td align="center"><code>PDtype</code></td>
<td align="left">(required only when \code{diversity = "PD"}), select PD type: </code>PDtype = "PD"</code> (effective total branch length) or </code>PDtype = "meanPD"</code> (effective number of equally divergent lineages). Default is </code>"meanPD"</code>, where </code>meanPD = PD/tree depth</code>.</td>
</tr>
<tr class="odd">
<td align="center"><code>FDdistM</code></td>
<td align="left">(required only when </code>diversity = "FD"</code>), a species pairwise distance matrix for all species in the pooled assemblage.</td>
</tr>
<tr class="even">
<td align="center"><code>FDtype</code></td>
<td align="left">(required only when </code>diversity = "FD"</code>), select FD type: </code>FDtype = "tau_values"</code> for FD under specified threshold values, or </code>FDtype = "AUC"</code> (area under the curve of tau-profile) for an overall FD which integrates all threshold values between zero and one. Default is </code>"AUC"</code>.</td>
</tr>
<tr class="odd">
<td align="center"><code>FDtau</code></td>
<td align="left">(required only when </code>diversity = "FD"</code> and </code>FDtype = "tau_values"</code>), a numerical vector between 0 and 1 specifying tau values (threshold levels). If </code>NULL</code> (default), then threshold is set to be the mean distance between any two individuals randomly selected from the pooled assemblage (i.e., quadratic entropy).</td>
</tr>


</tbody>
</table>

## DATA FORMAT/INFORMATION
Three types of data are supported:

1. Individual‐based abundance data (`datatype="abundance"`): Input data for each assemblage/site include samples species abundances in an empirical sample of n individuals (“reference sample”). When there are N assemblages, input data consist of an S by N abundance matrix, or N lists of species abundances.  

2. Sampling‐unit‐based incidence data: There are two kinds of input data.  
 (1) Incidence‐raw data (`datatype="incidence_raw"`): for each assemblage, input data for a reference sample consist of a species‐by‐sampling‐unit matrix; when there are N assemblages, input data consist of N lists of matrices, and each matrix is a species‐by‐sampling‐unit matrix. If the phylogenetic diversity required, the matrix of combined assemblage is allowed, but nT must be specified (see above description)..  
  (2) Incidence‐frequency data (`datatype="incidence_freq"`): input data for each assemblage consist of species sample incidence frequencies (row sums of each incidence matrix). When there are N assemblages, input data consist of an S by N matrix, or N lists of species incidence frequencies. The first entry of each list must be the total number of sampling units, followed by the species incidence frequencies.
  
## RAREFACTION/EXTRAPOLATION VIA EXAMPLES

Two data sets (beetles for abundance data ) are included in iNEXT3D package. The first list of beetles' data consist of abundance data who live on three treatments (“Control”, “Debarked” and "Scratched") of trees. The second list consist of the pylogenetic tree for every beetles. And the third list consist of distance matrix for each pair of beetles. For these data, the following commands display the sample species abundances and run the `iNEXT3D()` function for three types of diversty  (`"TD"`, `"PD"`, `"FD"` with threshold dmean, `"AUC"` which integate FD from threshold 0-1) in `q = 0`. 

```{r eval=FALSE}
data("beetles.abu")
str(beetles.abu$data)
iNEXT3D(beetles.abu$data, diversity = "TD", q = c(0, 1, 2), datatype = "abundance")
```


If one diversity class required, then `iNEXT3D()` function returns including: `$Info` for summarizing data information; `$iNextEst` for showing diversity estimates along with related statistics for a series of rarefied and extrapolated samples; and `$AsyEst` for showing asymptotic diversity estimates along with related statistics, otherwise, returns lists which length equal to the number of diversity class required and also named by diversity class. Among each list include three data frames:`$Info`, `$iNextEst` and `$AsyEst`. 

`$Info`, as shown below, returns basic data information.  It can also be presented using function `DataInfo()` (for "TD"), `PDInfo()` (for "PD") and `FDInfo()` (for "FD").
```
$TD$TDInfo: phylogenetic taxonomic basic data information
  Assembalge       n S.obs     SC  f1  f2 f3 f4 f5 f6 f7 f8 f9 f10
1    control   11346	  84  0.998  21  13  6  2  2  1  0  0  1   2
2   debarked    3201    61  0.994  19   7  3  1  0  1  2  2  0   2
3  scratched   10716    86  0.997  27   5  7  5  2  1  1  1  1   0
```

```{r eval=FALSE}
data("beetles")
str(beetles.abu$data)
iNEXT3D(beetles.abu$data, diversity = "PD", q = c(0, 1, 2), datatype = "abundance", PDtree = beetles.abu$tree)
```

```
$PD$PDInfo: phylogenetic basic data information
  Assemblage      n  S.obs    PD.obs  f1* f2*       g1        g2  Reftime
1   control   11346	    84  9000.431	21  14  2068.110	1281.121  276.437
2  debarked   22034     61  6584.298  19  10  2037.802   416.104	276.437
3 scratched   81991     86  8873.475  27   8  2189.268   875.041  276.437
```

```{r eval=FALSE}
data("beetles")
str(beetles.abu$data)
iNEXT3D(beetles.abu$data, diversity = "FD", q = c(0, 1, 2), datatype = "abundance", FDdistM = beetles.abu$distance.matrix)
```

```
$FD$FDInfo: functional basic data information
  Assembalge       n S.obs     SC  f1  f2 f3 f4 f5 f6 f7 f8 f9 f10
1    control   81940	  84  0.998   3   6  4  3  1  0  0  0  1   1
2   debarked    3201    61  0.994   8   1  1  1  2  1  0  0  0   2
3  scratched    3201    86  0.997   6   9  4  2  2  1  1  0  0   0
```

```{r eval=FALSE}
data("beetles")
str(beetles.abu$data)
iNEXT3D(beetles.abu$data, diversity = "AUC", q = c(0, 1, 2), datatype = "abundance", FDdistM = beetles.abu$distance.matrix)
```

```
$AUC$AUCInfo: phylogenetic taxonomic basic data information
  Assembalge       n S.obs     SC  
1    control   11346	  84  0.998  
2   debarked    3201    61  0.994  
3  scratched   10716    86  0.997  
```
<p>

Data description needed !!!. Diversity estimates and related statistics are computed for these 40 knots (corresponding to sample sizes m = 1, 598, 1195, …, 22692), which locates the reference sample at the mid‐point of the selected knots. If the argument se=TRUE, then the bootstrap method is applied to obtain the 95% confidence intervals for each diversity and sample coverage estimates. 

For the sample size corresponding to each knot, the first data frame of list `$iNextEst` (as shown below for 'size_based') under each diversity class includes the sample size (`m`, i.e., each of the 40 knots), the method (`interpolated`, `observed`, or `extrapolated`, depending on whether the size `m` is less than, equal to, or greater than the reference sample size), the diversity order, the diversity estimate of order q (`qD`, `qPD`, `qFD` and `AUC`), the 95% lower and upper confidence limits of diversity (`qD.LCL` with`qD.UCL`, `qPD.LCL` with `qPD.UCL`, `qFD.LCL` with `qFD.UCL` and `AUC.LCL` with `AUC.UCL`), and the sample coverage estimate (`SC`) along with the 95% lower and upper confidence limits of sample coverage (`SC.LCL`, `SC.UCL`). These sample coverage estimates with confidence intervals are used for plotting the sample completeness curve. Time consuming for `class = "AUC"`, therefore the demo shows only first three class.


```{r, echo=FALSE}
data("beetles")

out.TD <- iNEXT3D(data = beetles.abu$data, diversity = "TD", 
               q = c(0, 1, 2), datatype = "abundance", 
               nboot = 0)
out.PD <- iNEXT3D(data = beetles.abu$data, diversity = "PD", 
               q = c(0, 1, 2), datatype = "abundance", 
               PDtree = beetles.abu$tree, 
               nboot = 0)
out.FD <- iNEXT3D(data = beetles.abu$data, diversity = "FD", 
               q = c(0, 1, 2), datatype = "abundance", 
               FDdistM = beetles.abu$distance.matrix,
               nboot = 0)
#out$TD$DataInfo
```

```{r eval=FALSE}
head(out.TD$TDiNextEst$size_based)
head(out.PD$PDiNextEst$size_based)
head(out.FD$FDiNextEst$size_based)
```


```
$TD$TDiNextEst: taxonomic diversity estimates with rarefied and extrapolated samples.
$size_based
   Assemblage   m        Method Order.q     qD qD.LCL qD.UCL    SC SC.LCL SC.UCL Type
1  control      1    Rarefaction      0  1.000  1.000  1.000 0.224  0.216  0.231   TD
10 control     598   Rarefaction      0 37.188 36.501 37.874 0.982  0.981  0.983   TD
NOTE: Only show two estimates, call iNEXT.object$TD$iNextEst$size_based. to show complete output.
```

```
$PD$PDiNextEst: phylogenetic diversity estimates with rarefied and extrapolated samples.
$size_based
   Assemblage   m        Method Order.q     qPD   qPD.LCL  qPD.UCL    SC SC.LCL SC.UCL threshold  Type
1  control      1   Rarefaction      0  276.437  275.040  277.834 0.224  0.216  0.231   276.437    PD
10 control     598   Rarefaction     0 4127.571 4031.506 4223.636 0.982  0.981  0.983   276.437    PD
NOTE: Only show two estimates, call iNEXT.object$PD$iNextEst$size_based. to show complete output.
```
```
$FD$FDiNextEst: funcational diversity estimates with rarefied and extrapolated samples.
$size_based
   Assemblage   m        Method Order.q     qFD qFD.LCL qFD.UCL    SC SC.LCL SC.UCL  threshold  Type
1  control      1    Rarefaction      0   1.000   0.999   1.000 0.224  0.216  0.231      0.100    FD
10 control     598   Rarefaction      0  13.152  12.708  13.597 0.982  0.981  0.983      0.100    FD
NOTE: Only show two estimates, call iNEXT.object$FD$iNextEst$size_based. to show complete output.
```
The second data frame of list `$iNextEst` (as shown below for 'coverage_based') under each class includes the reference sample coverage ('goalSC'), real sample coverage ('SC'), sample size (`m`, i.e., each of the 40 knots), the method (`Rarefaction`, `Observed`, or `Extrapolation`, depending on whether the size `m` is less than, equal to, or greater than the reference sample size), the diversity order, the diversity estimate of order q (`qD`, `qPD`, `qFD` and `AUC`) conditioning on 'goalSC', the 95% lower and upper confidence limits of diversity (`qD.LCL` with `qD.UCL`, `qPD.LCL` with `qPD.UCL`, `qFD.LCL` with `qFD.UCL` and `AUC.LCL` with `AUC.UCL`). These sample coverage estimates with confidence intervals are used for plotting the coverage-based R/E curves. 


```{r eval=FALSE}
head(out.TD$TDiNextEst$coverage_based)
head(out.PD$PDiNextEst$coverage_based)
head(out.FD$FDiNextEst$coverage_based)
```

```
TD$TDiNextEst$coverage_based
   Assemblage goalSC    SC   m        Method order.q     qD qD.LCL qD.UCL Type
1  Girdled     0.224 0.224   1    Rarefaction      0  1.000  0.980  1.019   TD
10 Girdled     0.982 0.982  598   Rarefaction      0 37.188 35.800 38.575   TD
NOTE: Only show two estimates, call iNEXT.object$TD$iNextEst$coverage_based. to show complete output.
```
```
$PD$PDiNextEst$coverage_based
   Assemblage goalSC    SC   m        Method order.q      qPD  qpD.LCL  qpD.UCL Reftime Type
1  Girdled     0.224 0.224   1    Rarefaction      0  276.437  275.210  277.664 276.437   PD
10 Girdled     0.982 0.982  598   Rarefaction      0 4127.571 3871.915 4383.227 276.437   PD
NOTE: Only show two estimates, call iNEXT.object$PD$iNextEst$coverage_based. to show complete output.
```
```
$FD$FDiNextEst$coverage_based
   Assemblage goalSC    SC   m        Method order.q    qFD  qfD.LCL  qfD.UCL threshold Type
1  Girdled     0.224 0.224   1   Rarefaction       0  1.000   0.999    1.000      0.100   FD
10 Girdled     0.982 0.982  598  Rarefaction       0 13.152  12.170   14.134      0.100   FD
NOTE: Only show two estimates, call iNEXT.object$FD$iNextEst$coverage_based to show complete output.
```

```{r eval=FALSE}
out.TD$AsyEst
```

`$AsyEst` lists the observed diversity, asymptotic estimates, estimated bootstrap s.e. and 95% confidence intervals for Hill numbers with q = 0, 1, and 2. The estimated asymptotes are calculated via the functions `Asy3D()` and the observed diversity are calculated via the functions `Obs3D()`. The output for the beetles data is shown below. All row and column variables are self‐explanatory. 

```
$TD$TDAsyEst: asymptotic taxonomic diversity estimates along with related statistics.
Assemblage         Diversity  Observed Estimator     s.e.      LCL      UCL
   control  Species richness    84.000   100.960   12.313   84.000  125.094
           Shannon diversity     9.805     9.852    0.172    9.514   10.190
           Simpson diversity     4.457     4.458    0.076    4.309    4.608
NOTE: Only show one assemblage, call iNEXT.object$TD$AsyEst to show complete output.
```

```{r eval=FALSE}
out.PD$PDAsyEst
```

```
$PD$PDAsyEst: asymptotic phylogenetic diversity estimates along with related statistics.
                 Phylogenetic   Phylogenetic   Phylogenetic
Assemblage          diversity       Observed      Estimator       s.e.        LCL        UCL
   control   Species richness       9000.430      10669.554   1526.004   9000.430  13660.467
            Shannon diversity        816.638        818.103      6.180    805.989    830.217
            Simpson diversity        509.542        473.711      2.451    468.907    478.515
NOTE: Only show one assemblage, call iNEXT.object$PD$AsyEst to show complete output.
```

```{r eval=FALSE}
out.FD$FDAsyEst
```

```
$FD$FDAsyEst: asymptotic functional diversity estimates along with related statistics.
                   Functional   Functional    Functional
Assemblage          diversity     Observed     Estimator    s.e.     LCL     UCL
   control   Species richness       27.234        28.049   0.912  27.234  29.837
            Shannon diversity        4.304         4.309   0.046   4.218   4.400
            Simpson diversity        2.967         2.967   0.024   2.920   3.015
NOTE: Only show one assemblage, call iNEXT.object$FD$AsyEst to show complete output.
```


The user may specify an integer sample size for the argument endpoint to designate the maximum sample size of R/E calculation. For species richness, the extrapolation method is reliable up to the double reference sample size; beyond that, the prediction bias may be large. However, for measures of q = 1 and 2, the extrapolation can usually be safely extended to the asymptotic if data are not sparse; thus there is no limit for the value of endpoint for these two measures. 

The user may also specify the number of knots in the range of sample size between 1 and the endpoint. If you choose a large number of knots, then it may take a long time to obtain the output due to the time‐consuming bootstrap method. Alternatively, the user may specify a series of sample sizes for R/E computation, as in the following example:

```{r eval=FALSE}
# set a series of sample sizes (m) for R/E computation
m <- c(1, 5, 20, 50, 100, 200, 400)
iNEXT3D(beetles.abu[[1]], diversity = c("TD", "PD", "FD", "AUC"), 
       q = 0, datatype="abundance", size = m,
       tree = beetles.abu[[2]], 
       distM = beetles.abu[[3]])
```

Further, `iNEXT3D` can simultaneously run R/E computation for Hill numbers with q = 0, 1, and 2 by specifying a vector for the argument `q` as follows:

```{r eval=FALSE}
iNEXT3D(beetles.abu[[1]], diversity = c("TD", "PD", "FD", "AUC"),
        q = c(0,1,2), datatype = "abundance", size = m,
        tree = beetles.abu[[2]], 
        distM = beetles.abu[[3]])
```

## GRAPHIC DISPLAYS: FUNCTION ggiNEXT3D()

The function `ggiNEXT3D()`, which extends `ggplot2` with default arguments, is described as follows: 
```{r eval=FALSE}
ggiNEXT3D(outcome, type = 1:3, se = TRUE,
          facet.var = "Assemblage", color.var = "Order.q")  
```
Here `outcome` is the object of `iNEXT3D()`'s output. Three types of curves are allowed for different diversity classes:  

(1) Sample-size-based R/E curve (`type=1`): see Figs. 1a and 2a in the main text. This curve plots diversity estimates with confidence intervals (if `se=TRUE`) as a function of sample size up to double the reference sample size, by default, or a user‐specified `endpoint`.  

(2) Sample completeness curve (`type=2`) with confidence intervals (if `se=TRUE`): see Figs. 1b and 2b in the main text. This curve plots the sample coverage with respect to sample size for the same range described in (1).  

(3) Coverage-based R/E curve (`type=3`): see Figs. 1c and 2c in the main text. This curve plots the diversity estimates with confidence intervals (if `se=TRUE`) as a 
function of sample coverage up to the maximum coverage obtained from the maximum size described in (1). 

The argument `facet.var=("Order.q", "Assemblage")` is used to create a separate plot for each value of the specified variable. For example, the following code displays a separate plot (in Figs 1a and 1c) for each value of the diversity order q. The `ggiNEXT3D()` function is a wrapper around 
`ggplot2` package to create a R/E curve using a single line of code. The resulting object is of class `"ggplot"`, so can be manipulated using the `ggplot2` tools. 


The argument `facet.var="Assemblage"` in `ggiNEXT3D` function creates a separate plot for each assembalge, therefore the different Order.q will seperated by different colours as shown below: 
```{r }
# Sample‐size‐based R/E curves, separating by "assemblage""
ggiNEXT3D(out.TD, type = 1, facet.var = "Assemblage")


```

The argument `facet.var="Order.q"` in `ggiNEXT3D` function creates a separate plot for each order, therefore three assemblages will be seperated by different colours as shown below: 
```{r }
# Sample‐size‐based R/E curves, separating by "Order.q"
ggiNEXT3D(out.TD, type = 1, facet.var = "Order.q")


```


The following commands return the sample completeness curve in which different colors are used for the three assemblages. Since the sample completeness curve are same for differnet class of diversity, `ggiNEXT3D` returns only one plot: 

```{r}
ggiNEXT3D(out.TD, type = 2, facet.var = "Order.q", color.var="Assemblage")
```


The following commands return the coverage‐based R/E sampling curves in which different colors are used for the three assemblages (`facet.var="Assemblage"`) and for three orders 
(`facet.var="Order.q"`)

```{r}
ggiNEXT3D(out.TD, type = 3, facet.var="Assemblage")
```



```{r}
ggiNEXT3D(out.TD, type = 3, facet.var="Order.q")
```



## INCIDENCE DATA


For illustration, we use the Hinkley's fish data (in the dataset fish included in the package) at three time periods (2011-2013, 2014-2016 and 2017-2019). Data.frame of first list consist of incidence frequencies are shown below. The first row must be the total number of sampling units, followed by the species incidence frequencies.

```{r}
data("fish")
head(fish[[1]])

```

Note that incidence frequency data (`datatype = "incidence_freq`) is allowed only for diversity class: `"TD"`, `"FD"` and `"AUC"`.If `"PD"` required, use data form of incidence raw. The following commands return three time period R/E sampling curves for fish data. The argument `color.var =  "Assemblage"` is used to display curves in different colors for three time periods.

```{r}
data("fish")
out.inc <- iNEXT3D(data = fish$incidence_freq, 
                   diversity = "FD",
                   q = c(0,1,2), datatype = "incidence_freq", 
                   FDdistM = fish$dis)


# Sample‐size‐based R/E curves
ggiNEXT3D(out.inc, type = 1, color.var = "Assemblage")
```



```{r}
# Sample completeness curves
ggiNEXT3D(out.inc, type = 2, color.var = "Assemblage") 
```


```{r}
# Coverage‐based R/E curves
ggiNEXT3D(out.inc, type = 3, color.var = "Assemblage")     
```


## RAW INCIDENCE DATA FUNCTION: incidence_raw

Incidence raw data is allowed for all diversity class. The data set (`fish.raw`) included in the package is three list of matrices; each matrix is a species by plots data.frame for a time period. 

```{r}
data("fish.raw")
head(fish.raw[[1]])
head(fish.raw[[2]])
head(fish.raw[[3]])
```


```{r}
fish.tree = fish$tree
fish.dis = fish$dis

out.raw <- iNEXT3D(data = fish.raw, 
                   diversity = "PD",
                   q = c(0, 1, 2), datatype = "incidence_raw", 
                   PDtree = fish.tree)
ggiNEXT3D(out.raw, type = 1)
```



```{r}
out.raw <- iNEXT3D(data = fish.raw, 
                   diversity = "FD",
                   q = c(0, 1, 2), datatype = "incidence_raw", 
                   FDdistM = fish.dis)

ggiNEXT3D(out.raw, type = 1)
```

