---
title: "A Quick Introduction to iNEXT.3D via Examples"
author: "Anne Chao"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    vignette: >
      %\VignetteIndexEntry{A Quick Introduction to iNEXT.3D via Examples}
      %\VignetteEngine{knitr::rmarkdown}
      %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "", 
                      fig.retina=2,
                      fig.align='center',
                      fig.width = 7, fig.height = 5,
                      warning = FALSE, message = FALSE)
options("width"=200)

if("gridExtra" %in% rownames(installed.packages()) == FALSE) {install.packages("gridExtra")}
# devtools::load_all(".")
library(iNEXT.3D)
library(ggplot2)
# data("beetles.abu")
```


```{r eval=FALSE}
## install iNEXT.3D package from CRAN
# install.packages("iNEXT.3D")
## install the latest version from github
install.packages('devtools')
library(devtools)
# install_github('AnneChao/iNEXT.3D')
## import packages
library(iNEXT.3D)
library(ggplot2)
```


In this document, we provide a quick introduction demonstrating how to run the package `iNEXT.3D`(iNterpolation and EXTrapolation in three Dimensions). `iNEXT.3D` has several main function: `iNEXT3D`, `ggiNEXT3D`, `AO3D`, `ggAO3D`, `estimate3D`, and `DataInfo3D.`

`iNEXT3D` focuses on three measures of Hill numbers of order q: species richness (`q = 0`), Shannon diversity (`q = 1`, the exponential of Shannon entropy) and Simpson diversity (`q = 2`, the inverse of Simpson concentration) among three classes: taxonomic diversity ('TD'), phylogenetic diversity ('PD'), functional diversity ('FD'). For each diversity measure, `iNEXT.3D` uses the observed sample of abundance or incidence data (called the “reference sample”) to compute diversity estimates and the associated confidence intervals for the following two types of rarefaction and extrapolation (R/E): 

1. Sample‐size‐based R/E sampling curves:`iNEXT3D` computes diversity estimates for rarefied and extrapolated samples up to an appropriate size. This type of sampling curve plots the diversity estimates with respect to sample size.     
2. Coverage‐based R/E sampling curves: `iNEXT3D` computes diversity estimates for rarefied and extrapolated samples with sample completeness (as measured by sample coverage) up to an appropriate coverage. This type of sampling curve plots the diversity estimates with respect to sample coverage. 

`iNEXT3D` also plots the above two types of sampling curves and a sample completeness curve. The sample completeness curve provides a bridge between these two types of curves.

## MAIN FUNCTION: iNEXT3D()

We first describe the main function `iNEXT3D()` with default arguments: 

```{r eval=FALSE}
iNEXT3D(data, diversity = 'TD', q = c(0,1,2), datatype = "abundance", 
        size = NULL, endpoint = NULL, knots = 40, nboot = 50, conf = 0.95, nT = NULL, 
        PDtree = NULL, PDreftime = NULL, PDtype = 'meanPD', 
        FDdistM, FDtype = 'AUC', FDtau = NULL)
```

The arguments of this function are briefly described below, and will be explained in more details by illustrative examples in later text. This main function computes diversity estimates of order q, the sample coverage estimates and related statistics for K (if `knots=K`) evenly‐spaced knots (sample sizes) between size 1 and the `endpoint`, where the endpoint is described below. Each knot represents a particular sample size for which diversity estimates will be calculated. By default, endpoint = double the reference sample size (total sample size for abundance data; total sampling units for incidence data). For example, if `endpoint = 10`, `knot = 4`, diversity estimates will be computed for a sequence of samples with sizes (1, 4, 7, 10).  
<table style="width:100%;">
<colgroup>
<col width="20%">
<col width="80%">
</colgroup>
<thead>
<tr class="header">
<th align="center">Argument</th>
<th align="left">Description</th>
</tr>
</thead>
<tbody>
<tr class="odd">
<td align="center"><code>data</code></td>
<td align="left">
(a) For </code>datatype = "abundance"</code>, data can be input as a vector of species abundances (for a single assemblage), matrix/data.frame (species by assemblages), or a list of species abundance vectors. \cr
(b) For </code>datatype = "incidence_freq"</code>, data can be input as a vector of incidence frequencies (for a single assemblage), matrix/data.frame (species by assemblages), or a list of incidence frequencies; the first entry in all types of input must be the number of sampling units in each assemblage. \cr
(c) For </code>datatype = "incidence_raw"</code>, data can be input as a list of matrix/data.frame (species by sampling units); data can also be input as a matrix/data.frame by merging all sampling units across assemblages based on species identity; in this case, the number of sampling units (nT, see below) must be input. </td>
</tr>
<tr class="even">
<td align="center"><code>diversity</code></td>
<td align="left">selection of diversity type: </code>'TD'</code> = Taxonomic diversity, </code>'PD'</code> = Phylogenetic diversity, and </code>'FD'</code> = Functional diversity.</td>
</tr>
<tr class="odd">
<td align="center"><code>q</code></td>
<td align="left">a numerical vector specifying the diversity orders. Default is c(0, 1, 2).</td>
</tr>
<tr class="even">
<td align="center"><code>datatype</code></td>
<td align="left">data type of input data: individual-based abundance data (</code>datatype = "abundance"</code>), sampling-unit-based incidence frequencies data (</code>datatype = "incidence_freq"</code>), or species by sampling-units incidence matrix (</code>datatype = "incidence_raw"</code>) with all entries being 0 (non-detection) or 1 (detection).</td>
</tr>
<tr class="odd">
<td align="center"><code>size</code></td>
<td align="left">an integer vector of sample sizes (number of individuals or sampling units) for which diversity estimates will be computed.
If NULL, then diversity estimates will be computed for those sample sizes determined by the specified/default </code>endpoint</code> and </code>knots</code>.</td>
</tr>
<tr class="even">
<td align="center"><code>endpoint</code></td>
<td align="left">an integer specifying the sample size that is the </code>endpoint</code> for rarefaction/extrapolation.
If NULL, then </code>endpoint</code> </code>=</code> double reference sample size.</td>
</tr>
<tr class="odd">
<td align="center"><code>knots</code></td>
<td align="left">an integer specifying the number of equally-spaced </code>knots</code> (say K, default is 40) between size 1 and the </code>endpoint</code>;each knot represents a particular sample size for which diversity estimate will be calculated.
If the </code>endpoint</code> is smaller than the reference sample size, then </code>iNEXT3D()</code> computes only the rarefaction esimates for approximately K evenly spaced </code>knots</code>. 
If the </code>endpoint</code> is larger than the reference sample size, then </code>iNEXT3D()</code> computes rarefaction estimates for approximately K/2 evenly spaced </code>knots</code> between sample size 1 and the reference sample size, and computes extrapolation estimates for approximately K/2 evenly spaced </code>knots</code> between the reference sample size and the </code>endpoint</code>.</td>
</tr>
<tr class="even">
<td align="center"><code>nboot</code></td>
<td align="left">a positive integer specifying the number of bootstrap replications when assessing sampling uncertainty and constructing confidence intervals. Enter 0 to skip the bootstrap procedures. Default is 50.</td>
</tr>
<tr class="odd">
<td align="center"><code>conf</code></td>
<td align="left">a positive number < 1 specifying the level of confidence interval. Default is 0.95.</td>
</tr>
<tr class="even">
<td align="center"><code>nT</code></td>
<td align="left">(required only when </code>datatype = "incidence_raw"</code> and input data is matrix/data.frame) a vector of nonnegative integers specifying the number of sampling units in each assemblage. If assemblage names are not specified, then assemblages are automatically named as "assemblage1", "assemblage2",..., etc. </td>
</tr>
<tr class="odd">
<td align="center"><code>PDtree</code></td>
<td align="left">(required only when </code>diversity = "PD"</code>), a phylogenetic tree in Newick format for all observed species in the pooled assemblage.</td>
</tr>
<tr class="even">
<td align="center"><code>PDreftime</code></td>
<td align="left">(required only when </code>diversity = "PD"</code>), a vector of numerical values specifying reference times for PD. Default is </code>NULL</code> (i.e., the age of the root of PDtree).</td>
</tr>
<tr class="odd">
<td align="center"><code>PDtype</code></td>
<td align="left">(required only when \code{diversity = "PD"}), select PD type: </code>PDtype = "PD"</code> (effective total branch length) or </code>PDtype = "meanPD"</code> (effective number of equally divergent lineages). Default is </code>"meanPD"</code>, where </code>meanPD = PD/tree depth</code>.</td>
</tr>
<tr class="even">
<td align="center"><code>FDdistM</code></td>
<td align="left">(required only when </code>diversity = "FD"</code>), a species pairwise distance matrix for all species in the pooled assemblage.</td>
</tr>
<tr class="odd">
<td align="center"><code>FDtype</code></td>
<td align="left">(required only when </code>diversity = "FD"</code>), select FD type: </code>FDtype = "tau_values"</code> for FD under specified threshold values, or </code>FDtype = "AUC"</code> (area under the curve of tau-profile) for an overall FD which integrates all threshold values between zero and one. Default is </code>"AUC"</code>.</td>
</tr>
<tr class="even">
<td align="center"><code>FDtau</code></td>
<td align="left">(required only when </code>diversity = "FD"</code> and </code>FDtype = "tau_values"</code>), a numerical vector between 0 and 1 specifying tau values (threshold levels). If </code>NULL</code> (default), then threshold is set to be the mean distance between any two individuals randomly selected from the pooled assemblage (i.e., quadratic entropy).</td>
</tr>


</tbody>
</table>

## DATA FORMAT/INFORMATION
Three types of data are supported:

1. Individual‐based abundance data (`datatype="abundance"`): Input data for each assemblage/site include samples species abundances in an empirical sample of n individuals (“reference sample”). When there are N assemblages, input data consist of an S by N abundance matrix, or N lists of species abundances.  

2. Sampling‐unit‐based incidence data: There are two kinds of input data.  
 (1) Incidence‐raw data (`datatype="incidence_raw"`): for each assemblage, input data for a reference sample consist of a species‐by‐sampling‐unit matrix; when there are N assemblages, input data consist of N lists of matrices, and each matrix is a species‐by‐sampling‐unit matrix. If the phylogenetic diversity required, the matrix of combined assemblage is allowed, but nT must be specified (see above description)..  
  (2) Incidence‐frequency data (`datatype="incidence_freq"`): input data for each assemblage consist of species sample incidence frequencies (row sums of each incidence matrix). When there are N assemblages, input data consist of an S by N matrix, or N lists of species incidence frequencies. The first entry of each list must be the total number of sampling units, followed by the species incidence frequencies.
  
## RAREFACTION/EXTRAPOLATION VIA EXAMPLES

Two data sets (beetles for abundance data ) are included in iNEXT3D package. The first list of beetles' data consist of abundance data who live on three treatments (“Control”, “Debarked” and "Scratched") of trees. The second list consist of the pylogenetic tree for every beetles. And the third list consist of distance matrix for each pair of beetles. For these data, the following commands display the sample species abundances and run the `iNEXT3D()` function for three types of diversty  (`"TD"`, `"PD"`, `"FD"` with threshold dmean, `"AUC"` which integate FD from threshold 0-1) in `q = 0`. 

If one diversity class required, then `iNEXT3D()` function returns including: `$Info` for summarizing data information; `$iNextEst` for showing diversity estimates along with related statistics for a series of rarefied and extrapolated samples; and `$AsyEst` for showing asymptotic diversity estimates along with related statistics, otherwise, returns lists which length equal to the number of diversity class required and also named by diversity class. Among each list include three data frames:`$Info`, `$iNextEst` and `$AsyEst`. 

`$Info`, as shown below, returns basic data information.  It can also be presented using function `DataInfo()` (for "TD"), `PDInfo()` (for "PD") and `AUCInfo()` (for "FD").

```{r eval=FALSE}
data("dunes")
str(dunes$data)
iNEXT3D(dunes$data, diversity = "TD", q = c(0, 1, 2), datatype = "abundance")
```

```
$DataInfo: Taxonomic basic data information
Assemblage  n   S.obs   SC     f1    f2    f3    f4    f5    f6    f7    f8    f9   f10

1 EM      373    17   0.992     3     1     1     0     0     1     0     1     1     1
2 MO      1490   39   0.999     2     1     1     4     2     2     3     1     1     1
3 TR      1059   42   0.996     4     2     4     2     2     2     0     1     2     2
```

```{r eval=FALSE}
data("dunes")
str(dunes$data)
iNEXT3D(dunes$data, diversity = "PD", q = c(0, 1, 2), datatype = "abundance", PDtree = dunes$tree)
```

```
$PDInfo: phylogenetic basic data information
Assemblage     n    S.obs PD.obs `f1*` `f2*`    g1    g2    Reftime
1 EM           373    17  2660.   3     1     309     65       325
2 MO          1490    39  4783.   2     1     211     43.3     325
3 TR          1059    42  5303.   4     2     352     108      325
```

```{r eval=FALSE}
data("dunes")
str(dunes$data)
iNEXT3D(dunes$data, diversity = "FD", q = c(0, 1, 2), datatype = "abundance", FDdistM = dunes$dist)
```

```
$AUCInfo: functional basic data information
  Assemblage  n   S.obs   SC     f1    f2    f3    f4    f5    f6    f7    f8    f9   f10
1 EM      373    17   0.992     3     1     1     0     0     1     0     1     1     1
2 MO      1490   39   0.999     2     1     1     4     2     2     3     1     1     1
3 TR      1059   42   0.996     4     2     4     2     2     2     0     1     2     2
```

<p>

Data description needed !!!. Diversity estimates and related statistics are computed for these 40 knots (corresponding to sample sizes m = 1, 177, 373, …, 746), which locates the reference sample at the mid‐point of the selected knots. If the argument se=TRUE, then the bootstrap method is applied to obtain the 95% confidence intervals for each diversity and sample coverage estimates. 

For the sample size corresponding to each knot, the first data frame of list `$iNextEst` (as shown below for 'size_based') under each diversity class includes the sample size (`m`, i.e., each of the 40 knots), the method (`interpolated`, `observed`, or `extrapolated`, depending on whether the size `m` is less than, equal to, or greater than the reference sample size), the diversity order, the diversity estimate of order q (`qD`, `qPD`, `qFD` and `qAUC`), the 95% lower and upper confidence limits of diversity (`qD.LCL` with`qD.UCL`, `qPD.LCL` with `qPD.UCL`, `qFD.LCL` with `qFD.UCL` and `qAUC.LCL` with `qAUC.UCL`), and the sample coverage estimate (`SC`) along with the 95% lower and upper confidence limits of sample coverage (`SC.LCL`, `SC.UCL`). These sample coverage estimates with confidence intervals are used for plotting the sample completeness curve. Time consuming for `class = "AUC"`, therefore the demo shows only first three class.


```{r, echo=FALSE}
data("dunes")

out.TD <- iNEXT3D(data = dunes$data, diversity = "TD", 
               q = c(0, 1, 2), datatype = "abundance", 
               nboot = 10)
out.PD <- iNEXT3D(data = dunes$data, diversity = "PD", 
               q = c(0, 1, 2), datatype = "abundance", 
               PDtree = dunes$tree, 
               nboot = 10)
out.FD <- iNEXT3D(data = dunes$data, diversity = "FD", 
               q = c(0, 1, 2), datatype = "abundance", 
               FDdistM = dunes$dist,
               nboot = 5)
#out$TD$DataInfo
```

```{r}
head(out.TD$iNextEst$size_based)
head(out.PD$PDiNextEst$size_based)
head(out.FD$AUCiNextEst$size_based)
```


The second data frame of list `$iNextEst` (as shown below for 'coverage_based') under each class includes the reference sample coverage ('goalSC'), real sample coverage ('SC'), sample size (`m`, i.e., each of the 40 knots), the method (`Rarefaction`, `Observed`, or `Extrapolation`, depending on whether the size `m` is less than, equal to, or greater than the reference sample size), the diversity order, the diversity estimate of order q (`qD`, `qPD`, `qFD` and `qAUC`) conditioning on 'goalSC', the 95% lower and upper confidence limits of diversity (`qD.LCL` with `qD.UCL`, `qPD.LCL` with `qPD.UCL`, `qFD.LCL` with `qFD.UCL` and `qAUC.LCL` with `qAUC.UCL`). These sample coverage estimates with confidence intervals are used for plotting the coverage-based R/E curves. 


```{r}
head(out.TD$iNextEst$coverage_based)
head(out.PD$PDiNextEst$coverage_based)
head(out.FD$AUCiNextEst$coverage_based)
```


`$AsyEst` lists the observed diversity, asymptotic estimates, estimated bootstrap s.e. and 95% confidence intervals for Hill numbers with q = 0, 1, and 2. The estimated asymptotes and the observed diversity are calculated via the functions `AO3D()` . The output for the dunes data is shown below. All row and column variables are self‐explanatory. 

```{r}
out.TD$AsyEst
```

```{r}
out.PD$PDAsyEst
```

```{r}
out.FD$AUCAsyEst
```


The user may specify an integer sample size for the argument endpoint to designate the maximum sample size of R/E calculation. For species richness, the extrapolation method is reliable up to the double reference sample size; beyond that, the prediction bias may be large. However, for measures of q = 1 and 2, the extrapolation can usually be safely extended to the asymptotic if data are not sparse; thus there is no limit for the value of endpoint for these two measures. 

The user may also specify the number of knots in the range of sample size between 1 and the endpoint. If you choose a large number of knots, then it may take a long time to obtain the output due to the time‐consuming bootstrap method. Alternatively, the user may specify a series of sample sizes for R/E computation, as in the following example:

```{r eval=FALSE}
# set a series of sample sizes (m) for R/E computation
m <- c(1, 5, 20, 50, 100, 200, 400)
iNEXT3D(dunes$data, diversity = "TD", 
       q = 0, datatype="abundance", size = m)
```

Further, `iNEXT3D` can simultaneously run R/E computation for Hill numbers with q = 0, 1, and 2 by specifying a vector for the argument `q` as follows:

```{r eval=FALSE}
iNEXT3D(dunes$data, diversity = "TD",
        q = c(0,1,2), datatype = "abundance", size = m)
```

## GRAPHIC DISPLAYS: FUNCTION ggiNEXT3D()

The function `ggiNEXT3D()`, which extends `ggplot2` with default arguments, is described as follows: 
```{r eval=FALSE}
ggiNEXT3D(outcome, type = 1:3, se = TRUE,
          facet.var = "Assemblage", color.var = "Order.q")  
```
Here `outcome` is the object of `iNEXT3D()`'s output. Three types of curves are allowed for different diversity classes:  

(1) Sample-size-based R/E curve (`type=1`): see Figs. 1a and 2a in the main text. This curve plots diversity estimates with confidence intervals (if `se=TRUE`) as a function of sample size up to double the reference sample size, by default, or a user‐specified `endpoint`.  

(2) Sample completeness curve (`type=2`) with confidence intervals (if `se=TRUE`): see Figs. 1b and 2b in the main text. This curve plots the sample coverage with respect to sample size for the same range described in (1).  

(3) Coverage-based R/E curve (`type=3`): see Figs. 1c and 2c in the main text. This curve plots the diversity estimates with confidence intervals (if `se=TRUE`) as a 
function of sample coverage up to the maximum coverage obtained from the maximum size described in (1). 

The argument `facet.var=("Order.q", "Assemblage")` is used to create a separate plot for each value of the specified variable. For example, the following code displays a separate plot (in Figs 1a and 1c) for each value of the diversity order q. The `ggiNEXT3D()` function is a wrapper around 
`ggplot2` package to create a R/E curve using a single line of code. The resulting object is of class `"ggplot"`, so can be manipulated using the `ggplot2` tools. 


The argument `facet.var="Assemblage"` in `ggiNEXT3D` function creates a separate plot for each assembalge, therefore the different Order.q will seperated by different colours as shown below: 
```{r }
# Sample‐size‐based R/E curves, separating by "assemblage""
ggiNEXT3D(out.TD, type = 1, facet.var = "Assemblage")
```

The argument `facet.var="Order.q"` in `ggiNEXT3D` function creates a separate plot for each order, therefore three assemblages will be seperated by different colours as shown below: 
```{r }
# Sample‐size‐based R/E curves, separating by "Order.q"
ggiNEXT3D(out.TD, type = 1, facet.var = "Order.q")
```


The following commands return the sample completeness curve in which different colors are used for the three assemblages. Since the sample completeness curve are same for differnet class of diversity, `ggiNEXT3D` returns only one plot: 

```{r}
ggiNEXT3D(out.TD, type = 2, facet.var = "Order.q", color.var="Assemblage")
```


The following commands return the coverage‐based R/E sampling curves in which different colors are used for the three assemblages (`facet.var="Assemblage"`) and for three orders
(`facet.var="Order.q"`)

```{r}
ggiNEXT3D(out.TD, type = 3, facet.var="Assemblage")
```

```{r}
ggiNEXT3D(out.TD, type = 3, facet.var="Order.q")
```


## RAW INCIDENCE DATA FUNCTION: incidence_raw

For illustration, we use the Hinkley's fish data (in the dataset fish included in the package) at three time periods (1981-1985, 1987-1991 and 2015-2019).
Incidence raw data is allowed for all diversity class.This data set (`fish`) included in the package is three list of matrices; each matrix is a species by plots data.frame for a time period. We only use Taxonomic diversity(TD) for demonstration below.

```{r}
data("fish")
head(fish[[1]]$`1981-1985`,4)
```


```{r }
# fish.tree = fish$tree
# fish.dis = fish$dist

out.raw <- iNEXT3D(data = fish$data, 
                   diversity = "TD",
                   q = c(0, 1, 2), datatype = "incidence_raw",nboot = 10)
ggiNEXT3D(out.raw, type = 1)
```


```{r}
ggiNEXT3D(out.raw, type = 2)
```

```{r}
ggiNEXT3D(out.raw, type = 3)
```

## FREQUENCIES INCIDENCE DATA FUNCTION: incidence_freq

We transform Hinkley's fish data(incidence raw data) to incidence freqencies data(incidence_freq).
Data.frame of first list consist of incidence frequencies are shown below. The first row must be the total number of sampling units, followed by the species incidence frequencies.
We only use Taxonomic diversity(TD) for demonstration below.

```{r}
fish.freq = cbind(c(ncol(fish$data$`1981-1985`),rowSums(fish$data$`1981-1985`)),
                  c(ncol(fish$data$`1987-1991`),rowSums(fish$data$`1987-1991`)),
                  c(ncol(fish$data$`2015-2019`),rowSums(fish$data$`2015-2019`)))
rownames(fish.freq)[1]="sample units"
colnames(fish.freq)=names(fish$data)
```

```{r}
head(fish.freq)
```

Note that incidence frequency data (`datatype = "incidence_freq`) is allowed only for diversity class: `"TD"`, `"FD"`.If `"PD"` required, use data form of incidence raw. The following commands return three time period R/E sampling curves for fish data. The argument `color.var =  "Assemblage"` is used to display curves in different colors for three time periods.

```{r}
out.incfreq <- iNEXT3D(data = fish.freq, 
                   diversity = "TD",
                   q = c(0,1,2), datatype = "incidence_freq",nboot = 10)


# Sample‐size‐based R/E curves
ggiNEXT3D(out.incfreq, type = 1, color.var = "Assemblage")
```



```{r}
# Sample completeness curves
ggiNEXT3D(out.incfreq, type = 2, color.var = "Assemblage") 
```

```{r}
# Coverage‐based R/E curves
ggiNEXT3D(out.incfreq, type = 3, color.var = "Assemblage")     
```



## DATA INFORMATION FUNCTION: DataInfo3D()

We can supply the function
```{r eval=FALSE}
DataInfo3D(data, diversity = "TD", datatype = "abundance", 
  nT = NULL, PDtree, PDreftime = NULL, 
  FDdistM, FDtype = "AUC", FDtau = NULL) 
```
to compute three type diversity('TD','PD','FD') data information, which including sample size, observed species richness, sample coverage estimate, and the first ten abundance/incidence frequency counts, and so on.

```{r}
DataInfo3D(dunes$data, diversity = 'TD', datatype = "abundance")
```

## POINT ESTIMATION FUNCTION: estimate3D()
We also supply the function 
```{r eval=FALSE}
estimate3D(data, diversity = "TD", q = c(0, 1, 2), datatype = "abundance", 
  base = "coverage", level = NULL, nboot = 50, conf = 0.95, 
  nT = NULL, PDtree, PDreftime = NULL, PDtype = "meanPD", 
  FDdistM, FDtype = "AUC", FDtau = NULL) 
```

to compute three type diversity('TD','PD','FD') estimates with q = 0, 1, 2 for any particular level of sample size (`base="size"`) or any specified level of sample coverage (`base="coverage"`) for either abundance data (`datatype="abundance"`) or incidence data (`datatype="incidence_freq" or "incidence_raw"`). If `level=NULL`, this function computes the diversity estimates for the minimum sample size/coverage among all assemblages. 

For example, the following command returns the taxonomic diversity ('TD') with a specified level of sample coverage of 93% for the dunes data. For some assemblages, this coverage value corresponds to the rarefaction part whereas the others correspond to extrapolation, as indicated in the method of the output. 

```{r}
estimate3D(dunes$data, diversity = 'TD', q = c(0,1,2), datatype = "abundance", base = "coverage",level = 0.93)
```


## EMPIRICAL AND ASYMPTOTIC DIVERSITY FUNCTION: AO3D

```{r,eval=FALSE}
AO3D(
  data,diversity = "TD",q = seq(0,2,0.2),datatype = "abundance",
  nboot = 50,conf = 0.95,nT = NULL,method = c("Asymptotic", "Observed"),
  PDtree,PDreftime = NULL,PDtype = "meanPD",
  FDdistM,FDtype = "AUC",FDtau = NULL
)
```

The function `AO3D()` can compute three type diversity('TD','PD','FD'),which including empirical diversity and asymptotic diversity. For either abundance data (`datatype="abundance"`) or incidence data(`datatype="incidence_freq" or "incidence_raw"`)with any specified level of q can be compute.

For example, the following command returns an empirical taxonomic diversity('TD') and asymptotic taxonomic diversity('TD') for dunes data, along with its confidence interval, for a specified q level from 0 to 2.

```{r}
out1 <- AO3D(dunes$data, diversity = 'TD', datatype = "abundance", method = c("Asymptotic", "Observed"),nboot = 5,conf = 0.95)

out1
```


## GRAPHIC DISPLAYS FUNCTION: ggAO3D()

Plots q-profile, time-profile, and tau-profile based on the outcome of AO3D using the ggplot2 package.

The function `ggAO3D()`, which extends `ggplot2` with default arguments, is described as follows: 

```{r,eval=FALSE}
ggAO3D(outcome, profile = "q")
```
Here `outcome` is the object of `AO3D`'s output and `profile` is an object versus to diversity.Default is `profile = "q"`.Note that `profile = "time"` is allowed for only when diversity = "PD" and `profile = "tau"` profile is allowed for only when `diversity = "FD"` and `FDtype = "tau_values"`.  

```{r }
# q profile curve""
ggAO3D(out1,profile = "q")
```

The argument `profile = "time"` in `ggAO3D` function creates a separate plot for each order.q. Therefore the different assemblages will seperated by different colours as shown below: 
```{r }
# time curves, separating by "Order.q"
data(data.inc)
data <- data.inc$data
tree <- data.inc$tree
nT <- data.inc$nT
out2 <- AO3D(data, diversity = 'PD', q = c(0, 1, 2), datatype = "incidence_raw", nT = nT, nboot = 5, method = c("Asymptotic", "Observed"), PDtree = tree, PDreftime = seq(0.1, 82.8575, length.out = 40))
ggAO3D(out2, profile = "time")
```

The argument `profile = "tau"` in `ggAO3D` function creates a separate plot for each order.q. Therefore the different assemblages will seperated by different colours as shown below: 
```{r }
# tau curves, separating by "Order.q"
data <- dunes$data
distM <-  dunes$dist
out3 <- AO3D(data, diversity = 'FD', q = c(0, 1, 2), datatype = "abundance", nboot = 5, method = c("Asymptotic", "Observed"), FDtau = seq(0, 1, 0.1), FDdistM = distM, FDtype = 'tau_values')
ggAO3D(out3, profile = "tau")
```

