---
title: "A Quick Introduction to iNEXT.3D via Examples"
author: "Anne Chao, K.H. Hu,"
date: "`r Sys.Date()`"
output:
  rmarkdown::html_vignette:
    vignette: >
      %\VignetteIndexEntry{A Quick Introduction to iNEXT.3D via Examples}
      %\VignetteEngine{knitr::rmarkdown}
      %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(collapse = TRUE, comment = "", 
                      fig.retina=2,
                      fig.align='center',
                      fig.width = 7, fig.height = 5,
                      warning = FALSE, message = FALSE)
options("width"=200)

if("gridExtra" %in% rownames(installed.packages()) == FALSE) {install.packages("gridExtra")}
# devtools::load_all(".")
library(iNEXT.3D)
library(ggplot2)
# data("beetles.abu")
```


```{r eval=FALSE}
## install iNEXT.3D package from CRAN
# install.packages("iNEXT.3D")  # coming soon

## install the latest version from github
install.packages('devtools')
library(devtools)
install_github('AnneChao/iNEXT.3D')

## import packages
library(iNEXT.3D)
library(ggplot2)
```


In this document, here provide a quick introduction demonstrating how to run the package `iNEXT.3D` (iNterpolation and EXTrapolation in three Dimensions). `iNEXT.3D` has several main functions: `iNEXT3D`, `ggiNEXT3D`, `AO3D`, `ggAO3D`, `estimate3D`, and `DataInfo3D.`

`iNEXT.3D` is the extension of R package [iNEXT](https://cran.r-project.org/web/packages/iNEXT/index.html) (Hsieh et al., 2016). `iNEXT.3D` focuses on three measures of Hill numbers of order q: species richness (`q = 0`), Shannon diversity (`q = 1`, the exponential of Shannon entropy) and Simpson diversity (`q = 2`, the inverse of Simpson concentration) and extend Hill number to three dimensions: taxonomic diversity (TD), phylogenetic diversity (PD), and functional diversity (FD) under Hill-Chao family frame work (Chao et al., 2019). Besides, `iNEXT.3D` also provide statistic estimation for three dimensions biodiversity (Chao et al., 2021). For each diversity measure, `iNEXT.3D` uses the observed sample of abundance or incidence data (called the “reference sample”) to compute diversity estimates and the associated confidence intervals for the following two types of rarefaction and extrapolation (R/E): 

1. Sample-size-based R/E sampling curves:`iNEXT3D` computes diversity estimates for rarefied and extrapolated samples up to an appropriate size. This type of sampling curve plots the diversity estimates with respect to sample size.     
2. Coverage-based R/E sampling curves: `iNEXT3D` computes diversity estimates for rarefied and extrapolated samples with sample completeness (as measured by sample coverage) up to an appropriate coverage. This type of sampling curve plots the diversity estimates with respect to sample coverage. 

`iNEXT.3D` also plots the above two types of sampling curves and a sample completeness curve by `ggiNEXT3D`. The sample completeness curve provides a bridge between these two types of curves.

## How to cite
If you publish your work based on results from iNEXT.3D package, you should make references to the following Online reference:

*	Chao, A., Henderson, P. A., Chiu, C.-H., Moyes, F., Hu, K.-H., Dornelas, M and. Magurran, A. E. (2021). Measuring temporal change in alpha diversity: a framework integrating taxonomic, phylogenetic and functional diversity and the iNEXT.3D standardization. Methods in Ecology and Evolution, 12, 1926-1940.

## SOFTWARE NEEDED TO RUN iNEXT.3D IN R
- Required: [R](https://cran.r-project.org/)
- Suggested: [RStudio IDE](https://www.rstudio.com/products/RStudio/#Desktop)

## HOW TO RUN iNEXT.3D:
The `iNEXT.3D` package can be downloaded from Anne Chao's [iNEXT.3D_github](https://github.com/AnneChao/iNEXT.3D) using the following commands. For a first-time installation, an additional visualization extension package (`ggplot2`) must be installed and loaded. 

```{r eval=FALSE}
## install iNEXT.3D package from CRAN
# install.packages("iNEXT.3D")  # coming soon

## install the latest version from github
install.packages('devtools')
library(devtools)
install_github('AnneChao/iNEXT.3D')

## import packages
library(iNEXT.3D)
library(ggplot2)
```

An online version of iNEXT.3D(https://chao.shinyapps.io/iNEXT_3D/) is also available for users without R background.


## MAIN FUNCTION: iNEXT3D()

First describe the main function `iNEXT3D()` with default arguments: 

```{r eval=FALSE}
iNEXT3D(data, diversity = 'TD', q = c(0,1,2), datatype = "abundance", 
        size = NULL, endpoint = NULL, knots = 40, nboot = 50, conf = 0.95, nT = NULL, 
        PDtree = NULL, PDreftime = NULL, PDtype = 'meanPD', 
        FDdistM, FDtype = 'AUC', FDtau = NULL)
```

The arguments of this function are briefly described below, and will be explained in more details by illustrative examples in later text. This main function computes diversity estimates of order q, the sample coverage estimates and related statistics for K (if `knots = K`) evenly-spaced knots (sample sizes) between size 1 and the `endpoint`, where the endpoint is described below. Each knot represents a particular sample size for which diversity estimates will be calculated. By default, `endpoint` = double the reference sample size for abundance data or double the total sampling units for incidence data. For example, if `endpoint = 10`, `knot = 4`, diversity estimates will be computed for a sequence of samples with sizes (1, 4, 7, 10).  
<table style="width:100%;">
<colgroup>
<col width="20%">
<col width="80%">
</colgroup>
<thead>
<tr class="header">
<th align="center">Argument</th>
<th align="left">Description</th>
</tr>
</thead>
<tbody>
<tr class="odd">
<td align="center"><code>data</code></td>
<td align="left">
(a) For <code>datatype = "abundance"</code>, data can be input as a vector of species abundances (for a single assemblage), matrix/data.frame (species by assemblages), or a list of species abundance vectors. \cr
(b) For <code>datatype = "incidence_freq"</code>, data can be input as a vector of incidence frequencies (for a single assemblage), matrix/data.frame (species by assemblages), or a list of incidence frequencies; the first entry in all types of input must be the number of sampling units in each assemblage. \cr
(c) For <code>datatype = "incidence_raw"</code>, data can be input as a list of matrix/data.frame (species by sampling units); data can also be input as a matrix/data.frame by merging all sampling units across assemblages based on species identity; in this case, the number of sampling units (<code>nT</code>, see below) must be input. </td>
</tr>
<tr class="even">
<td align="center"><code>diversity</code></td>
<td align="left">selection of diversity type: <code>'TD'</code> = Taxonomic diversity, <code>'PD'</code> = Phylogenetic diversity, and <code>'FD'</code> = Functional diversity.</td>
</tr>
<tr class="odd">
<td align="center"><code>q</code></td>
<td align="left">a numerical vector specifying the diversity orders. Default is <code>c(0, 1, 2)</code>.</td>
</tr>
<tr class="even">
<td align="center"><code>datatype</code></td>
<td align="left">data type of input data: individual-based abundance data (<code>datatype = "abundance"</code>), sampling-unit-based incidence frequencies data (<code>datatype = "incidence_freq"</code>), or species by sampling-units incidence matrix (<code>datatype = "incidence_raw"</code>) with all entries being 0 (non-detection) or 1 (detection).</td>
</tr>
<tr class="odd">
<td align="center"><code>size</code></td>
<td align="left">an integer vector of sample sizes (number of individuals or sampling units) for which diversity estimates will be computed.
If <code>NULL</code>, then diversity estimates will be computed for those sample sizes determined by the specified/default <code>endpoint</code> and <code>knots</code>.</td>
</tr>
<tr class="even">
<td align="center"><code>endpoint</code></td>
<td align="left">an integer specifying the sample size that is the <code>endpoint</code> for rarefaction/extrapolation.
If NULL, then <code>endpoint</code> <code>=</code> double reference sample size.</td>
</tr>
<tr class="odd">
<td align="center"><code>knots</code></td>
<td align="left">an integer specifying the number of equally-spaced <code>knots</code> (say K, default is 40) between size 1 and the <code>endpoint</code>;each knot represents a particular sample size for which diversity estimate will be calculated.
If the <code>endpoint</code> is smaller than the reference sample size, then <code>iNEXT3D()</code> computes only the rarefaction esimates for approximately K evenly spaced <code>knots</code>. 
If the <code>endpoint</code> is larger than the reference sample size, then <code>iNEXT3D()</code> computes rarefaction estimates for approximately K/2 evenly spaced <code>knots</code> between sample size 1 and the reference sample size, and computes extrapolation estimates for approximately K/2 evenly spaced <code>knots</code> between the reference sample size and the <code>endpoint</code>.</td>
</tr>
<tr class="even">
<td align="center"><code>nboot</code></td>
<td align="left">a positive integer specifying the number of bootstrap replications when assessing sampling uncertainty and constructing confidence intervals. Enter 0 to skip the bootstrap procedures. Default is 50.</td>
</tr>
<tr class="odd">
<td align="center"><code>conf</code></td>
<td align="left">a positive number < 1 specifying the level of confidence interval. Default is 0.95.</td>
</tr>
<tr class="even">
<td align="center"><code>nT</code></td>
<td align="left">(required only when <code>datatype = "incidence_raw"</code> and input data is matrix/data.frame) a vector of nonnegative integers specifying the number of sampling units in each assemblage. If assemblage names are not specified, then assemblages are automatically named as "assemblage1", "assemblage2",..., etc. </td>
</tr>
<tr class="odd">
<td align="center"><code>PDtree</code></td>
<td align="left">(required only when <code>diversity = "PD"</code>), a phylogenetic tree in Newick format for all observed species in the pooled assemblage.</td>
</tr>
<tr class="even">
<td align="center"><code>PDreftime</code></td>
<td align="left">(required only when <code>diversity = "PD"</code>), a vector of numerical values specifying reference times for PD. Default is <code>NULL</code> (i.e., the age of the root of PDtree).</td>
</tr>
<tr class="odd">
<td align="center"><code>PDtype</code></td>
<td align="left">(required only when <code>diversity = "PD"</code>), select PD type: <code>PDtype = "PD"</code> (effective total branch length) or <code>PDtype = "meanPD"</code> (effective number of equally divergent lineages). Default is <code>"meanPD"</code>, where <code>meanPD = PD/tree depth</code>.</td>
</tr>
<tr class="even">
<td align="center"><code>FDdistM</code></td>
<td align="left">(required only when <code>diversity = "FD"</code>), a species pairwise distance matrix for all species in the pooled assemblage.</td>
</tr>
<tr class="odd">
<td align="center"><code>FDtype</code></td>
<td align="left">(required only when <code>diversity = "FD"</code>), select FD type: <code>FDtype = "tau_values"</code> for FD under specified threshold values, or <code>FDtype = "AUC"</code> (area under the curve of tau-profile) for an overall FD which integrates all threshold values between zero and one. Default is <code>"AUC"</code>.</td>
</tr>
<tr class="even">
<td align="center"><code>FDtau</code></td>
<td align="left">(required only when <code>diversity = "FD"</code> and <code>FDtype = "tau_values"</code>), a numerical vector between 0 and 1 specifying tau values (threshold levels). If <code>NULL</code> (default), then threshold is set to be the mean distance between any two individuals randomly selected from the pooled assemblage (i.e., quadratic entropy).</td>
</tr>


</tbody>
</table>

## DATA FORMAT/INFORMATION
Three types of data are supported:

1. Individual-based abundance data (`datatype = "abundance"`): Input data for each assemblage/site include samples species abundances in an empirical sample of n individuals (“reference sample”). When there are N assemblages, input data consist of an S by N abundance matrix, or N lists of species abundances.  

2. Sampling-unit-based incidence data: There are two kinds of input data.  
 (a) Incidence-raw data (`datatype = "incidence_raw"`): for each assemblage, input data for a reference sample consist of a species-by-sampling-unit matrix; when there are N assemblages, input data consist of N lists of matrices, and each matrix is a species-by-sampling-unit matrix. The matrix of combined assemblage is allowed, but nT must be specified (see above description).
 
 (b) Incidence-frequency data (`datatype = "incidence_freq"`): input data for each assemblage consist of species sample incidence frequencies (row sums of each incidence matrix). When there are N assemblages, input data consist of an (S + 1) by N matrix, or N lists of species incidence frequencies. The first entry of each list must be the total number of sampling units, followed by the species incidence frequencies.
  
## RAREFACTION/EXTRAPOLATION VIA EXAMPLES

The abundance data set, dunes, is included in `iNEXT.3D` package. The first list of dunes data is consists of abundance data with three landscape(“EM”, “MO” and "TR"). The second list is consist of the pylogenetic tree for all species. And the third list is consist of distance matrix for each pair of species. For this data set, the following commands display the abundance data and run the `iNEXT3D()` function for three dimensions of diversity(`TD`, `PD`, `FD`) under order `q = 0, 1, 2`. 

The output of function `iNEXT3D()` includes three parts: `$DataInfo` for summarizing data information; `$iNextEst` for showing diversity estimates along with related statistics for a series of rarefied and extrapolated samples; and `$AsyEst` for showing asymptotic diversity estimates along with related statistics under taxonomic diversity. Result under phylogenetic diversity or functional diversity includes these three parts, too. 

`$DataInfo`, as shown below, returns basic data information.  It can also be presented by the function `DataInfo3D()`.

```{r eval=FALSE}
data("dunes")
str(dunes$data)
out.TD = iNEXT3D(dunes$data, diversity = "TD", q = c(0, 1, 2), datatype = "abundance")
out.TD[1]
```

```
$DataInfo: Taxonomic basic data information
Assemblage  n   S.obs   SC     f1    f2    f3    f4    f5    f6    f7    f8    f9   f10

1 EM      373    17   0.992     3     1     1     0     0     1     0     1     1     1
2 MO      1490   39   0.999     2     1     1     4     2     2     3     1     1     1
3 TR      1059   42   0.996     4     2     4     2     2     2     0     1     2     2
```

```{r eval=FALSE}
data("dunes")
str(dunes$data)
out.PD = iNEXT3D(dunes$data, diversity = "PD", q = c(0, 1, 2), 
                 datatype = "abundance", PDtree = dunes$tree)
out.PD[1]
```

```
$PDInfo: phylogenetic basic data information
Assemblage     n    S.obs PD.obs `f1*` `f2*`    g1    g2    Reftime
1 EM           373    17  2660.   3     1     309     65       325
2 MO          1490    39  4783.   2     1     211     43.3     325
3 TR          1059    42  5303.   4     2     352     108      325
```

```{r eval=FALSE}
data("dunes")
str(dunes$data)
out.FD = iNEXT3D(dunes$data, diversity = "FD", q = c(0, 1, 2), 
                 datatype = "abundance", FDdistM = dunes$dist, FDtype = "tau_values")
out.FD[1]
```

```
$FDInfo: functional basic data information
  Assemblage    n   S.obs   SC     a1'   a2'   a3'   a4'   a5'   a6'   a7'  a8'   a9'  a10'    Tau
1 EM          373    17   0.992     0     0     1     0     0     0     1     0     0     0  0.592
2 MO         1490    39   0.999     0     0     0     0     0     0     0     0     0     1  0.592
3 TR         1059    42   0.996     0     0     0     0     0     0     0     1     0     0  0.592
```

Second part of output from function `iNEXT3D` is diversity estimates and related statistics computed for these 40 knots by default (for example in "EM" assemblage, corresponding to sample sizes m = 1, 20, 40, ..., 372, 373, 374, ..., 746), which locates the reference sample size at the mid-point of the selected knots. The diversity can be based on sample-size-based and sample coverage-based. The first data frame of list `$iNextEst` (as shown below for 'size_based') includes the sample size (`m`), the `Method` (`Rarefaction`, `Observed`, or `Extrapolation`, depending on whether the size `m` is less than, equal to, or greater than the reference sample size), the diversity order (`Order.q`), the diversity estimate of order q (`qD` in TD, `qPD` in PD, `qFD` in FD (under specified thresholds), `qAUC` in FD (area under curve)), the lower and upper confidence limits of diversity (`qD.LCL` and `qD.UCL` in TD, `qPD.LCL` and `qPD.UCL` in PD, `qFD.LCL` and `qFD.UCL` in FD (under specified thresholds), `qAUC.LCL` and `qAUC.UCL` in FD (area under curve)) conditioning on sample size, and the sample coverage estimate (`SC`) along with the lower and upper confidence limits of sample coverage (`SC.LCL`, `SC.UCL`). These sample coverage estimates with confidence intervals are used for plotting the sample completeness curve. It is time consuming for `diversity = FD` and `FDtype = "AUC"`. If the argument `nboot` is greater than zero, then the bootstrap method is applied to obtain the confidence intervals for each diversity and sample coverage estimates. 

Here only show first six rows for each diversity:

```{r, echo=FALSE}
data("dunes")

out.TD <- iNEXT3D(data = dunes$data, diversity = "TD", q = c(0, 1, 2), 
                  datatype = "abundance", nboot = 30)
out.PD <- iNEXT3D(data = dunes$data, diversity = "PD", q = c(0, 1, 2), 
                  datatype = "abundance", PDtree = dunes$tree, nboot = 30)
out.FD <- iNEXT3D(data = dunes$data, diversity = "FD", q = c(0, 1, 2), 
                  datatype = "abundance", FDdistM = dunes$dist, FDtype = "tau_values", nboot = 30)

head(out.TD$iNextEst$size_based)
head(out.PD$PDiNextEst$size_based)
head(out.FD$FDiNextEst$size_based)
```

```{r eval=FALSE}
out.TD$iNextEst$size_based
out.PD$PDiNextEst$size_based
out.FD$FDiNextEst$size_based
```


The second data frame of list `$iNextEst` (as shown below for 'coverage_based') includes the sample coverage estimate ('SC'), the sample size (`m`), the `Method` (`Rarefaction`, `Observed`, or `Extrapolation`, depending on whether the size `m` is less than, equal to, or greater than the reference sample size), the diversity order (`Order.q`), the diversity estimate of order q (`qD` in TD, `qPD` in PD, `qFD` in FD (under specified thresholds), `qAUC` in FD (area under curve)), the lower and upper confidence limits of diversity (`qD.LCL` and `qD.UCL` in TD, `qPD.LCL` and `qPD.UCL` in PD, `qFD.LCL` and `qFD.UCL` in FD (under specified thresholds), `qAUC.LCL` and `qAUC.UCL` in FD (area under curve)) conditioning on sample coverage estimate. 

Here only show first six rows for each diversity:

```{r, echo=FALSE}
head(out.TD$iNextEst$coverage_based)
head(out.PD$PDiNextEst$coverage_based)
head(out.FD$FDiNextEst$coverage_based)
```

```{r, eval=FALSE}
out.TD$iNextEst$coverage_based
out.PD$PDiNextEst$coverage_based
out.FD$FDiNextEst$coverage_based
```

The output `$AsyEst` lists the diversity labels (`Diversity` in TD, `Phylogenetic Diversity` in PD, `Functional Diversity` in FD), the observed diversity (`Observed` in TD, `Phylogenetic Observed` in PD, `Functional Observed` in FD), asymptotic diversity estimates  (`Estimator` in TD, `Phylogenetic Estimator` in PD, `Functional Estimator` in FD), estimated bootstrap standard error (`s.e.`) and  confidence intervals for diversity with q = 0, 1, and 2 (`LCL`, `UCL`). The estimated asymptotic and observed diversity can also be computed via the function `AO3D()`. The output are shown below:

```{r}
out.TD$AsyEst
```

```{r}
out.PD$PDAsyEst
```

```{r}
out.FD$FDAsyEst
```


The user can key in an integer sample size vector to designate the maximum sample size of R/E calculation. For species richness (order q = 0), the extrapolation method is reliable up to the double reference sample size; beyond that, the prediction bias may be large. However, for the measures of order q = 1 and 2, the extrapolation can reliably be extended to the asymptotic value (sample coverage = 1) if data are not sparse. If you choose a large number for bootstrap (`nboot`), it may take a long time to obtain the output. Following example below (don't show the output):

```{r eval=FALSE}
# set a sample size vector (m) for R/E computation
m <- c(1, 5, 20, 50, 100, 200, 400)
iNEXT3D(dunes$data, diversity = "TD", 
        q = 0, datatype="abundance", size = m)
```


## GRAPHIC DISPLAYS: FUNCTION ggiNEXT3D()

The function `ggiNEXT3D()`, which extends `ggplot2` with default arguments, is described as follows: 

```{r eval=FALSE}
ggiNEXT3D(outcome, type = 1:3, facet.var = "Assemblage", color.var = "Order.q")  
```

Here `outcome` is the object of `iNEXT3D()`'s output. Three types of curves are allowed for different diversity dimensions:  

(1) Sample-size-based R/E curve (`type = 1`): This curve plots diversity estimates with confidence intervals as a function of sample size.  

(2) Sample completeness curve (`type = 2`): This curve plots the sample coverage with respect to sample size. 

(3) Coverage-based R/E curve (`type = 3`): This curve plots the diversity estimates with confidence intervals as a function of sample coverage. 

The argument `facet.var = "Order.q"` or `facet.var = "Assemblage"` is used to create a separate plot for each value of the specified variable. For example, the following code displays a separate plot of the diversity order q. The `ggiNEXT3D()` function is a wrapper with package `ggplot2` to create a R/E curve in a single line of code. The figure object is of class `"ggplot"`, so can be manipulated by using the `ggplot2` tools. 

When `facet.var = "Assemblage"` in `ggiNEXT3D` function, it creates a separate plot for each assemblage and the different color lines represent each diversity order. Sample-size-based R/E curve (`type = 1`) as below:

```{r }
# Sample-size-based R/E curves, separating by "assemblage""
ggiNEXT3D(out.TD, type = 1, facet.var = "Assemblage")
```

When `facet.var = "Order.q"` in `ggiNEXT3D` function, it creates a separate plot for each diversity order and the different color lines represent each assemblage. Sample-size-based R/E curve (`type = 1`) as below:

```{r }
# Sample-size-based R/E curves, separating by "Order.q"
ggiNEXT3D(out.TD, type = 1, facet.var = "Order.q")
```


The following command return the sample completeness (sample coverage) curve (`type = 2`) in which different colors are used for the three assemblages. 

```{r}
ggiNEXT3D(out.TD, type = 2, facet.var = "Order.q", color.var = "Assemblage")
```


The following commands return the coverage-based R/E sampling curves in which different colors are used for the three assemblages (`facet.var = "Assemblage"`) and for three diversity orders (`facet.var = "Order.q"`).

```{r}
ggiNEXT3D(out.TD, type = 3, facet.var = "Assemblage")
```

```{r}
ggiNEXT3D(out.TD, type = 3, facet.var = "Order.q")
```


### EXAMPLE for INCIDENCE-RAW DATA

Incidence raw data is allowed for three diversity dimensions. For illustration, use the Hinkley's fish data (the dataset is included in the package) at three time periods (1981-1985, 1987-1991 and 2015-2019). This data set (`fish`) includes a list with three matrices; each matrix is a species-by-sampling-unit data. Here only use taxonomic diversity (TD) as demonstration below.

```{r}
data("fish")
head(fish$data$`1981-1985`)
```


```{r }
# fish.tree = fish$tree  ## for PD
# fish.dis = fish$dist   ## for FD

out.raw <- iNEXT3D(data = fish$data, diversity = "TD",
                   q = c(0, 1, 2), datatype = "incidence_raw", nboot = 30)
ggiNEXT3D(out.raw, type = 1)
```


```{r}
ggiNEXT3D(out.raw, type = 2)
```

```{r}
ggiNEXT3D(out.raw, type = 3)
```

### EXAMPLE for INCIDENCE-FREQUENCY DATA

We transform Hinkley's fish data (incidence-raw data) to incidence-frequency data (`incidence_freq`). The first row must be the total number of sampling units, followed by the species incidence frequencies in each assemblage. Here only use taxonomic diversity (TD) for demonstration below.

```{r}
fish.freq = cbind(c( ncol(fish$data$`1981-1985`), rowSums(fish$data$`1981-1985`) ),
                  c( ncol(fish$data$`1987-1991`), rowSums(fish$data$`1987-1991`) ),
                  c( ncol(fish$data$`2015-2019`), rowSums(fish$data$`2015-2019`) ))
rownames(fish.freq)[1] = "sample units"
colnames(fish.freq) = names(fish$data)
```

```{r}
head(fish.freq)
```

Note that incidence-frequency data (`datatype = "incidence_freq`) is allowed only for diversity class: `"TD"`, `"FD"`. If `"PD"` required, use incidence-raw data instead. The following commands return three R/E sampling curves for fish data. The argument `color.var =  "Order.q"` is used to display curves in different colors for diversity order.

```{r}
out.incfreq <- iNEXT3D(data = fish.freq, diversity = "TD",
                       q = c(0, 1, 2), datatype = "incidence_freq",nboot = 30)


# Sample-size-based R/E curves
ggiNEXT3D(out.incfreq, type = 1, color.var = "Order.q")
```



```{r}
# Sample completeness curves
ggiNEXT3D(out.incfreq, type = 2)
```

```{r}
# Coverage-based R/E curves
ggiNEXT3D(out.incfreq, type = 3, color.var = "Order.q")     
```


## DATA INFORMATION FUNCTION: DataInfo3D()

```{r eval=FALSE}
DataInfo3D(data, diversity = "TD", datatype = "abundance", 
  nT = NULL, PDtree, PDreftime = NULL, 
  FDdistM, FDtype = "AUC", FDtau = NULL) 
```

Here provide the function `DataInfo3D` to compute three diversity dimensions ('TD', 'PD', 'FD') data information, which including sample size, observed species richness, sample coverage estimate, and the first ten abundance/incidence frequency counts when `diversity = TD`. And so on for PD, FD.

```{r}
DataInfo3D(dunes$data, diversity = 'TD', datatype = "abundance")
```

## POINT ESTIMATION FUNCTION: estimate3D()

```{r eval=FALSE}
estimate3D(data, diversity = "TD", q = c(0, 1, 2), datatype = "abundance", 
           base = "coverage", level = NULL, nboot = 50, conf = 0.95, 
           nT = NULL, PDtree, PDreftime = NULL, PDtype = "meanPD", 
           FDdistM, FDtype = "AUC", FDtau = NULL) 
```

`estimate3D` is used to compute three diversity dimensions (TD, PD, FD) estimates with q = 0, 1, 2 under any specified level of sample size (when `base = "size"`) or sample coverage (when `base = "coverage"`) for either abundance data (`datatype = "abundance"`) or incidence data (`datatype = "incidence_freq"` or `"incidence_raw"`). If `level = NULL`, this function computes the diversity estimates for the minimum sample size among all samples extrapolated to double reference sizes (when `base = "size"`) or the minimum sample coverage among all samples extrapolated to double reference sizes (when `base = "coverage"`). 

For example, the following command returns the taxonomic diversity ('TD') with a specified level of sample coverage = 99.5% for the dunes data. For some assemblages, this coverage value corresponds to the rarefaction part whereas the others correspond to extrapolation. 

```{r}
estimate3D(dunes$data, diversity = 'TD', q = c(0,1,2), datatype = "abundance", 
           base = "coverage", level = 0.995)
```


## EMPIRICAL AND ASYMPTOTIC DIVERSITY FUNCTION: AO3D

```{r,eval=FALSE}
AO3D(
  data, diversity = "TD", q = seq(0, 2, 0.2), datatype = "abundance",
  nboot = 50, conf = 0.95, nT = NULL, method = c("Asymptotic", "Observed"),
  PDtree, PDreftime = NULL, PDtype = "meanPD",
  FDdistM, FDtype = "AUC", FDtau = NULL
)
```

The function `AO3D()` compute three diversity dimensions (TD, PD, FD) for empirical (observed) diversity and estimated asymptotic diversity with any diversity order. For example, the following commands returns empirical and asymptotic taxonomic diversity ('TD') for dunes data, along with its confidence interval at diversity order q from 0 to 2. Here only show the first ten rows.

```{r}
out1 <- AO3D(dunes$data, diversity = 'TD', datatype = "abundance", 
             method = c("Asymptotic", "Observed"), nboot = 30, conf = 0.95)

head(out1, 10)
```


## GRAPHIC DISPLAYS FUNCTION: ggAO3D()

```{r,eval=FALSE}
ggAO3D(outcome, profile = "q")
```

`ggAO3D` plots q-profile, time-profile, and tau-profile based on `ggplot2`. Here `outcome` is the object from the function `AO3D`, and `profile` is a profile selection versus to diversity. Default is `profile = "q"`. Note that `profile = "time"` is allowed for only when diversity = "PD" and `profile = "tau"` profile is allowed for only when `diversity = "FD"` and `FDtype = "tau_values"`.  

```{r }
# q-profile curves
ggAO3D(out1,profile = "q")
```

The argument `profile = "time"` in `ggAO3D` function creates a separate plot for each diversity order q. Therefore the different assemblages will be represented by different color lines.

```{r }
# time-profile curves, separating by "Order.q"
data(data.inc)
data <- data.inc$data
tree <- data.inc$tree
nT <- data.inc$nT

out2 <- AO3D(data, diversity = 'PD', q = c(0, 1, 2), datatype = "incidence_raw", 
             nT = nT, nboot = 30, method = c("Asymptotic", "Observed"), 
             PDtree = tree, PDreftime = seq(0.1, 82.8575, length.out = 40))
ggAO3D(out2, profile = "time")
```

The argument `profile = "tau"` in `ggAO3D` function creates a separate plot for each diversity order q. Therefore the different assemblages will be represented by different color lines.

```{r }
# tau-profile curves, separating by "Order.q"
data <- dunes$data
distM <- dunes$dist

out3 <- AO3D(data, diversity = 'FD', q = c(0, 1, 2), datatype = "abundance", nboot = 30, method = c("Asymptotic", "Observed"), 
             FDtau = seq(0, 1, 0.1), FDdistM = distM, FDtype = 'tau_values')
ggAO3D(out3, profile = "tau")
```


## License
The iNEXT.3D package is licensed under the GPLv3. To help refine `iNEXT.3D`, your comments or feedback would be welcome (please send them to Anne Chao or report an issue on the iNEXT.3D github [iNEXT.3D_github](https://github.com/AnneChao/iNEXT.3D). 

## References
- Chao, A., Chiu, C.-H., Villéger, S., Sun, I.-F., Thorn, S., Lin, Y.-C., Chiang, J. M. and Sherwin, W. B. (2019). An attribute-diversity approach to functional diversity, functional beta diversity, and related (dis)similarity measures. Ecological Monographs, 89, e01343. 10.1002/ecm.1343.

- Chao, A., Henderson, P. A., Chiu, C.-H., Moyes, F., Hu, K.-H., Dornelas, M and Magurran, A. E. (2021). Measuring temporal change in alpha diversity: a framework integrating taxonomic, phylogenetic and functional diversity and the iNEXT.3D standardization. Methods in Ecology and Evolution, 12, 1926-1940.

- T.C. Hsieh, K. H. Ma, and Chao, A. (2016). iNEXT: An R package for rarefaction and extrapolation of species diversity (Hill numbers). Methods in Ecology and Evolution, 7, 1451-1456.


